Precise Data Flow Analysis\\in the Presence of Correlated Method Calls

### 2014-12-17 content/abstract.tex
pass through abstract, introduction, example, conclusion
spell check

\begin{abstract}
    When two methods are invoked on the same object, the dispatch behaviors of these
    method calls will be correlated. If two correlated method calls are polymorphic 
    (i.e., they dispatch to different method definitions depending on the type of 
    the receiver object), a program's interprocedural control flow graph (ICFG)
    will contain infeasible paths. Existing algorithms for dataflow analysis 
    are unable to ignore such infeasible paths, giving rise to loss of precision.

	We show how infeasible paths due to correlated calls can be eliminated for 
	\textit{Interprocedural Finite Distributive Subset} (IFDS) problems, a large class 
	of data-flow analysis problems with broad applications. Our approach is to
	transform an IFDS problem into an \textit{Interprocedural Distributive 
	Environment} (IDE) problem, in which edge functions are designed in a
	way that enables us to ignore dataflow along infeasible paths. A solution
	to this IDE problem can be mapped back to the solution space of the original
	IFDS problem. We prove the correctness of our approach, and report on
	some preliminary experiments with our prototype implementation. 
\end{abstract}

>>>>

\begin{abstract}
    When two methods are invoked on the same object, the dispatch behaviours of these
    method calls will be correlated. If two correlated method calls are polymorphic 
    (i.e., they dispatch to different method definitions depending on the type of 
    the receiver object), a program's interprocedural control flow graph (ICFG)
    will contain infeasible paths. Existing algorithms for dataflow analysis 
    are unable to ignore such infeasible paths, giving rise to loss of precision.

	We show how infeasible paths due to correlated calls can be eliminated for 
	\textit{Interprocedural Finite Distributive Subset} (IFDS) problems, a large class 
	of data-flow analysis problems with broad applications. Our approach is to
	transform an IFDS problem into an \textit{Interprocedural Distributive 
	Environment} (IDE) problem, in which edge functions are designed in a
	way that enables us to ignore dataflow along infeasible paths. A solution
	to this IDE problem can be mapped back to the solution space of the original
	IFDS problem.
        We formalize the approach, prove it correct, and report on an implementation
        in the WALA analysis framework.
\end{abstract}

### 2014-12-17 content/background.tex
remove completed comment
spell check

\section{Background}\label{sec:bg}
%The purpose of the correlated-calls analysis is to solve IFDS problems more precisely than using the standard IFDS algorithm by ruling out some infeasible paths. The correlated-calls analysis works by transforming an IFDS problem to an IDE problem, solving the IDE problem, and transforming the IDE result to a solution to the original IFDS problem. 
This section defines basic terminology and presents the IFDS and IDE algorithms.
%\ftodo{
  %Shall we add a sentence saying that readers familiar with these algorithms may skip this section?
%}

@@

The input to the IFDS algorithm is specified as
%Formally, a data-flow analysis problem is suitable for an IFDS analysis if it can be encoded as an IFDS problem
  $(G^*,\,D,\,F,\,M_F,\,\sqcap)$,
where $G^*=(N^*,\,E^*)$ is the supergraph of the input program with nodes $N^*$ and edges $E^*$,
$D$ is a finite set of \textit{data-flow facts},
$F$ is a set of distributive dataflow functions of type $2^D\to2^D$,
$M_F:\,E^*\to F$ assigns a dataflow function to each supergraph edge,
and $\sqcap$ is the \textit{meet operator} on the powerset $2^D$, either union or intersection.
In our presentation, the meet operator will always be union, but all of the results apply dually when the
meet is intersection.

@@

    To eliminate infeasible paths due to correlated calls, we encode the taint
    analysis using environments $\textsf{Env}(D,L)$, where $D$ is the set of
    variables and $L$ is a map from receiver variables to sets of possible
    types. The interpretation of such an environment $e$ is that a given variable
    $v \in D$ may contain a secret value in an exeuction in which the runtime
    types of the objects pointed to be the receiver variables are in the sets
    specified by $e(v)$.
\end{example}

@@

\ftodo{Would it make sense to discuss an example of an IDE problem here?  Perhaps we can do this
  without using much space by referring to an existing figure elsewhere in the paper?}
  \otodo{See Example 4 on the previous page.}

>>>>

\section{IFDS and IDE}\label{sec:bg}
%The purpose of the correlated-calls analysis is to solve IFDS problems more precisely than using the standard IFDS algorithm by ruling out some infeasible paths. The correlated-calls analysis works by transforming an IFDS problem to an IDE problem, solving the IDE problem, and transforming the IDE result to a solution to the original IFDS problem. 
This section defines basic terminology and presents the IFDS and IDE algorithms.
%\ftodo{
  %Shall we add a sentence saying that readers familiar with these algorithms may skip this section?
%}

@@

The input to the IFDS algorithm is specified as
%Formally, a data-flow analysis problem is suitable for an IFDS analysis if it can be encoded as an IFDS problem
  $(G^*,\,D,\,F,\,M_F,\,\sqcap)$,
where $G^*=(N^*,\,E^*)$ is the supergraph of the input program with nodes $N^*$ and edges $E^*$,
$D$ is a finite set of \textit{data-flow facts},
$F$ is a set of distributive dataflow functions of type $2^D\to2^D$,
$M_F:\,E^*\to F$ assigns a dataflow function to each supergraph edge,
and $\sqcap$ is the \textit{meet operator} on the power set $2^D$, either union or intersection.
In our presentation, the meet operator will always be union, but all of the results apply dually when the
meet is intersection.

@@

    To eliminate infeasible paths due to correlated calls, we encode the taint
    analysis using environments $\textsf{Env}(D,L)$, where $D$ is the set of
    variables and $L$ is a map from receiver variables to sets of possible
    types. The interpretation of such an environment $e$ is that a given variable
    $v \in D$ may contain a secret value in an execution in which the runtime
    types of the objects pointed to be the receiver variables are in the sets
    specified by $e(v)$.
\end{example}

@@

\commentout{
\ftodo{Would it make sense to discuss an example of an IDE problem here?  Perhaps we can do this
  without using much space by referring to an existing figure elsewhere in the paper?}
  \otodo{See Example 4 on the previous page.}
  \otodo{Now done in Example 5.}
}

### 2014-12-17 content/example.tex
spell check

As we discussed previously, the calls to \code{foo()} and \code{bar()} may dispatch
to different implementations of these methods in classes \code{A} and \code{B}, 
because the variable \code{a} that serves as the receiver for the call may be bound to
 objects of type \code{A} or type \code{B} at run time. 
However, consider the fact that the methods \code{foo()} and \code{bar()} are invoked
on \textit{the same object}. This means that the behaviors of the two method calls
are \textit{correlated}: if the call to \code{foo()} dispatches to \code{A.foo()},
then the call to \code{bar()} must dispatch to \code{A.bar()}, and if the call
to \code{foo()} dispatches to \code{B.foo()}, then the call to \code{bar()} must 
dispatch to \code{B.bar()}.  Consequently, paths such as the one highlighted in
Figure~\ref{fig:examplesupergraph} where the call to \code{foo()} dispatches 
to \code{A.foo()} and where the call to \code{bar()} dispatches to \code{B.bar()}
are infeasible.

@@

\subsection{An Example from the Scala Collections Library}
The Scala collections library contains the trait \code{TraversableOnce} that is
shared by both collections and iterators over them. 
\ftodo{
  I think it would be helpful to show a little more of the code of the trait here,
  such as the trait declaration and the signature of the toArray method containing the
  correlated calls. You can abbreviate everything else with ...
}
The \code{toArray} method of this trait
creates an array and copies the contents of the collection or iterator into it:
\inputMinted{java}{collectionsExample.java}
%\otodo{Marianna, can you figure out how to get LaTeX to syntax-highlight this?} 
When \code{this} refers to an iterator rather than a collection, the call to \code{this.size}
extracts all elements of the iterator to count them.
\ftodo{
  Would it make sense to show the code for the size method, and refer to the line numbers
  of that method in the above sentence?  
}
 At the call to \code{copyToArray},
the iterator is already empty, so nothing is copied to the newly created array.
\ftodo{
Would it make sense to show the code for copyToArray here?
}
One could design an IFDS analysis to detect this kind of bug.

However, the implementation of \code{TraversableOnce.toArray} is actually correct because the above
code is guarded with a test:
\inputMinted{java}{isTraversable.java}
\ftodo{
  It is confusing that this code fragment is labeled with line number 19 although it presumably
  occurs in the code before the previously discussed two lines. I think it is better to show
  all of the code at once, in a figure, and include at least the headers of all the methods mentioned
  here.
}
When the \code{isTraversableAgain} method returns false, as it does for
an iterator, the \code{toArray} method uses a different
(less efficient) implementation. The bug report would therefore be a false positive. 
The \code{isTraversableAgain} method is easy to analyze: it returns the constant
true in a collection and the constant false in an iterator. However, in order
to eliminate the false positive bug report, an analysis would need to rule out infeasible
paths using correlated calls. Specifically, the following path triggers the bug, but
is infeasible: first, call \code{isTraversableAgain} on a collection, returning
true, then call \code{size} and \code{copyToArray} on an iterator. Our correlated
calls analysis could determine that this path is infeasible because it calls the
collection version of \code{isTraversableAgain} but the iterator versions of
\code{size} and \code{copyToArray}.
\ftodo{Would it be helpful to create a diagram of the relevant fragments of the supergraph,
with the infeasible path in bold?}

>>>>

As we discussed previously, the calls to \code{foo()} and \code{bar()} may dispatch
to different implementations of these methods in classes \code{A} and \code{B}, 
because the variable \code{a} that serves as the receiver for the call may be bound to
 objects of type \code{A} or type \code{B} at run time. 
However, consider the fact that the methods \code{foo()} and \code{bar()} are invoked
on \textit{the same object}. This means that the behaviours of the two method calls
are \textit{correlated}: if the call to \code{foo()} dispatches to \code{A.foo()},
then the call to \code{bar()} must dispatch to \code{A.bar()}, and if the call
to \code{foo()} dispatches to \code{B.foo()}, then the call to \code{bar()} must 
dispatch to \code{B.bar()}.  Consequently, paths such as the one highlighted in
Figure~\ref{fig:examplesupergraph} where the call to \code{foo()} dispatches 
to \code{A.foo()} and where the call to \code{bar()} dispatches to \code{B.bar()}
are infeasible.

@@

\subsection{An Example from the Scala Collections Library}
The Scala collections library contains the trait \code{TraversableOnce} that is
shared by both collections and iterators over them. 
\ftodo{
  I think it would be helpful to show a little more of the code of the trait here,
  such as the trait declaration and the signature of the toArray method containing the
  correlated calls. You can abbreviate everything else with ...
}
The \code{toArray} method of this trait
creates an array and copies the contents of the collection or iterator into it:
\inputMinted{scala}{collectionsExample.scala}
%\otodo{Marianna, can you figure out how to get LaTeX to syntax-highlight this?} 
When \code{this} refers to an iterator rather than a collection, the call to \code{this.size}
extracts all elements of the iterator to count them.
\ftodo{
  Would it make sense to show the code for the size method, and refer to the line numbers
  of that method in the above sentence?  
}
 At the call to \code{copyToArray},
the iterator is already empty, so nothing is copied to the newly created array.
\ftodo{
Would it make sense to show the code for copyToArray here?
}
One could design an IFDS analysis to detect this kind of bug.

However, the implementation of \code{TraversableOnce.toArray} is actually correct because the above
code is guarded with a test:
\inputMinted{scala}{isTraversable.scala}
\ftodo{
  It is confusing that this code fragment is labeled with line number 19 although it presumably
  occurs in the code before the previously discussed two lines. I think it is better to show
  all of the code at once, in a figure, and include at least the headers of all the methods mentioned
  here.
}
When the \code{isTraversableAgain} method returns false, as it does for
an iterator, the \code{toArray} method uses a different
(less efficient) implementation. The bug report would therefore be a false positive. 
The \code{isTraversableAgain} method is easy to analyze: it returns the constant
true in a collection and the constant false in an iterator. However, in order
to eliminate the false positive bug report, an analysis would need to rule out infeasible
paths using correlated calls. Specifically, the following path triggers the bug, but
is infeasible: first, call \code{isTraversableAgain} on a collection, returning
true, then call \code{size} and \code{copyToArray} on an iterator. Our correlated
calls analysis could determine that this path is infeasible because it calls the
collection version of \code{isTraversableAgain} but the iterator versions of
\code{size} and \code{copyToArray}.
\ftodo{Would it be helpful to create a diagram of the relevant fragments of the supergraph,
with the infeasible path in bold?}

### 2014-12-17 content/intro.tex
spell check

This paper presents an approach to dataflow analysis that avoids a type
of infeasible path that arises in object-oriented programs when two or more
methods are dynamically dispatched on the same receiver object.
%\mtodo{Should we define the term ``receiver''?}
In such cases, if the
method calls are polymorphic (i.e., if they dispatch to different method definitions 
depending on the type of the receiver expression at run time), then their dispatch 
behaviors will be correlated. A recent paper \cite{DBLP:journals/scp/Tip15} identified
this problem but did not present a concrete solution or algorithm, and we are not aware 
of any existing dataflow analysis that is capable of avoiding infeasible paths that arise 
in the presence of correlated method calls.

>>>>

This paper presents an approach to dataflow analysis that avoids a type
of infeasible path that arises in object-oriented programs when two or more
methods are dynamically dispatched on the same receiver object.
%\mtodo{Should we define the term ``receiver''?}
In such cases, if the
method calls are polymorphic (i.e., if they dispatch to different method definitions 
depending on the type of the receiver expression at run time), then their dispatch 
behaviours will be correlated. A recent paper \cite{DBLP:journals/scp/Tip15} identified
this problem but did not present a concrete solution or algorithm, and we are not aware 
of any existing dataflow analysis that is capable of avoiding infeasible paths that arise 
in the presence of correlated method calls.
